---
title: "Criminology Data Prep"
author: "Ben Stockton"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = 'hold', warning = FALSE)
library(dplyr, warn.conflicts = FALSE)

# Suppress summarise info
options(dplyr.summarise.inform = FALSE)
library(magrittr)
library(ggplot2)
library(RColorBrewer)
library(mice)
library(lattice)
library(VIM)
library(haven)
library(readr)
library(cowplot)
source("helpers.R")
```


## Data Prep and Tidying

```{r}
dat <- read_dta("Data/Most Serious Sentence (2010-2019, limited).dta")
colnames(dat)
```

```{r, echo = FALSE}
dat$TRIAL <- factor(dat$TRIAL, labels = c("Yes", "Plea"))
dat$MALE <- factor(dat$MALE, labels = c("Female", "Male"))
dat$RECMIN <- factor(dat$RECMIN, labels = c("Yes", "No"))
```


```{r, echo = FALSE}
dat[dat$UNKNOWN == 1, "OFF_RACER"] <- NA
dat$OFF_RACER <- factor(dat$OFF_RACER, labels = c("WHITE", "BLACK", "LATINO", "OTHER"))
```

### Overview of the Predictors

Either tables or summaries with histograms of the predictors.

```{r, echo = FALSE, results = 'hide'}
table(dat$INCAR, useNA = "ifany")
table(dat$MALE, useNA = "ifany")
table(dat$BLACK, dat$LATINO, useNA = "ifany")
table(dat$OFF_RACER, useNA = "ifany")
table(dat$PERSONS, useNA = "ifany")
table(dat$RECMIN, useNA = "ifany")
table(dat$TRIAL, useNA = "ifany")
table(dat$PRS_0, useNA = "ifany")
table(dat$PRS_123, useNA = "ifany")
table(dat$PRS_45, useNA = "ifany")
table(dat$PRS_RR, useNA = "ifany")
```

Defendant Age at Trial

```{r}
hist(dat$DOSAGE, main = "Defendant Age at Trial", xlab = "Age")
summary(dat$DOSAGE)
```

Offense Gravity Score

```{r}
table(dat$OGS)
hist(dat$OGS, breaks = 16, main = "Offense Gravity Score", xlab = "Score")
summary(dat$OGS)
```

Previous Criminal History

```{r}
dat$PRS <- factor(dat$PRS_0 + 2 * dat$PRS_123 + 3 * dat$PRS_45 + 4 * dat$PRS_RR,
                  labels = c("None", "1/2/3", "4/5", "REVOC/RFEL"))
table(dat$PRS, useNA = "ifany")
```

Crime Type



```{r}
dat$CRIMETYPE <- factor(dat$PERSONS + 2 * dat$PROPERTY + 3 * dat$DRUGNEW 
                        + 4 * dat$DUI + 5 * dat$OTHER,
                        labels = c("PERSONS", "PROPERTY", "DRUGNEW",
                                   "DUI", "OTHER"))
table(dat$CRIMETYPE, useNA = "ifany")
```

Year


```{r}
dat$YEAR <- factor(dat$YEAR1 + 2 * dat$YEAR2 + 3 * dat$YEAR3 + 4 * dat$YEAR4 
                   + 5 * dat$YEAR5 + 6 * dat$YEAR6 + 7 * dat$YEAR7 
                   + 8 * dat$YEAR8 + 9 * dat$YEAR9 + 10 * dat$YEAR10 + 2009)
table(dat$YEAR, useNA = "ifany")
```

County


```{r}
dat$COUNTY <- rep(0, nrow(dat))
for (k in 1:67) {
    dat[dat[,paste0("COUNTY", k)] == 1, "COUNTY"] <- k
}
dat$COUNTY <- factor(dat$COUNTY)
table(dat$COUNTY, useNA = "ifany")
table(dat$COUNTY32, useNA = "ifany")
```

```{r}
dat.sm <- dat %>% select(INCAR, CRIMETYPE, OGS, OGSQ, RECMIN, TRIAL, PRS, 
                         MALE, DOSAGE, DOSAGEQ, OFF_RACER, COUNTY, YEAR)

# write_csv(dat.sm, "Data/most_serious_sentence_2010-2019_slim.csv")
rm(dat)

head(dat.sm)
```

## Missing Data Pattern

Checking the missing data pattern.

```{r, fig.width=8}
md.pattern(dat.sm[,c("INCAR", "OFF_RACER", "CRIMETYPE", "TRIAL", "OGS",
                     "RECMIN", "PRS", "DOSAGE", "MALE", "COUNTY", "YEAR")],
           rotate.names = T)
```

```{r}
c(834548, 26206, 2078, 34, 1465, 92, 3) / 86442
prop_mis <- (30005) / nrow(dat.sm)
```

`r 100 * round(prop_mis, 3)`% of the observations have at least one missing value. All missing values are in PRS, AGE, RECMIN, and RACE.

<!-- ```{r} -->
<!-- # marginplot(dat.sm[,c("RECMIN", "PRS")]) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- pbox(dat.sm, pos = 3, main = "Offense Gravity Score by Missingness") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- pbox(dat.sm, pos = 9, main = "Offender Age by Missingness") -->
<!-- ``` -->

## Logistic Regression with CCA on the full observed data

```{r}
# CCA Logistic Regression INCAR ~ .
fit <- glm(INCAR ~ ., data = dat.sm,
           family = binomial(link = "logit"))
# summary(fit)
coef(fit) %>% round(3)
beta <- coef(fit)
save(fit, file = "full_data_cca_log_reg.rda")
```

```{r}
rm(fit)
```

The odds ratio of the Race Effect (comparing Black defendants to White defendants) is `r round(exp(beta)["OFF_RACERBLACK"], 3)`. 

## Investigating Race by County

```{r}
race_by_county <- dat.sm %>% group_by(COUNTY) %>%
    summarise(prop_white = mean(OFF_RACER == "WHITE", na.rm = T),
              prop_black = mean(OFF_RACER == "BLACK", na.rm = T),
              prop_latino = mean(OFF_RACER == "LATINO", na.rm = T),
              prop_other = mean(OFF_RACER == "OTHER", na.rm = T)) %>% as.data.frame()
race_by_county %>% arrange(desc(prop_black)) %>% head(10)
race_by_county %>% arrange(desc(prop_white)) %>% head(10)
```

## Simulating Data 

I'll create a simulated "population" data set of the same size as the original data. For the simulations, I'll draw a small sample from the data.

For each predictor, I'll create a random sample from a distribution corresponding to the data type. So for a binary variable, I'll use $\hat{p}$ as the parameter of a Bernoulli distribution. For categorical and ordinal data, I'll draw randomly from the categories weighted by the observed prevalences. For continuous data I'll draw from a normal or gamma distribution depending on the shape and constraints on the real data. Finally the incarceration response is defined to be exactly $P(Y_i^{sim} = 1 | X^{sim}) = logit^{-1}(X^{sim}\hat{\beta})$.

```{r}
N.sim <- nrow(dat.sm)
dat.sim <- data.frame(INCAR = rep(0, N.sim))
```


### Simulating predictors

#### CRIMETYPE

Crimetype (Categorical, 5 levels) - Person, Property, Drug-new, DUI, Other

```{r}
pt <- prop.table(table(dat.sm$CRIMETYPE))

dat.sim$CRIMETYPE <- factor(sample(1:5, size = N.sim, prob = pt, replace = T),
                            labels = levels(dat.sm$CRIMETYPE))
p1 <- my_gg_bar(dat.sm, "CRIMETYPE")
p2 <- my_gg_bar(dat.sim, "CRIMETYPE")
cowplot::plot_grid(p1, p2, nrow = 1)
```

#### OGS

Offense Gravity Score (ordinal, 1-14)

```{r}
pt <- prop.table(table(dat.sm$OGS))

dat.sim$OGS <- sample(1:14, size = N.sim, prob = pt, replace = T)

p1 <- ggplot(dat.sm, aes(OGS)) +
    geom_histogram(bins = 14)
p2 <- ggplot(dat.sim, aes(OGS)) +
    geom_histogram(bins = 14)
cowplot::plot_grid(p1, p2, nrow = 1)
```


#### TRIAL 

Trial (Binary) - Yes = 1 or Plea Negotiation = 0

```{r}
pt <- prop.table(table(dat.sm$TRIAL))

dat.sim$TRIAL <- factor(sample(0:1, size = N.sim, prob = pt, replace = T), 
                        labels = c("Yes", "Plea"))
p1 <- my_gg_bar(dat.sm, "TRIAL")
p2 <- my_gg_bar(dat.sim, "TRIAL")
cowplot::plot_grid(p1, p2, nrow = 1)
```

#### MALE 

Gender (Binary) - Male = 1, Female = 0

```{r}
pt <- prop.table(table(dat.sm$MALE))

dat.sim$MALE <- factor(sample(0:1, size = N.sim, prob = pt, replace = T), 
                        labels = c("Female", "Male"))
p1 <- my_gg_bar(dat.sm, "MALE")
p2 <- my_gg_bar(dat.sim, "MALE")
cowplot::plot_grid(p1, p2, nrow = 1)
```


#### COUNTY (Categorical, 67 levels) - Counties in Pennsylvania

This variable should be viewed as a random effect in a properly set up model. Will need to decide if that's the direction I want to go. Keeping things as simple as possible would probably require selecting just a single county for analysis.

```{r}
pt <- prop.table(table(dat.sm$COUNTY))

dat.sim$COUNTY <- factor(sample(1:67, size = N.sim, prob = pt, replace = T), 
                        labels = levels(dat.sm$COUNTY))
p1 <- my_gg_bar(dat.sm, "COUNTY")
p2 <- my_gg_bar(dat.sim, "COUNTY")
cowplot::plot_grid(p1, p2, nrow = 2)
```

#### OFF_RACER 

Defendent Race (Categorical, 5 levels) - White (ref), Black, Latino, AAPI/Other, Unknown/NA 

Previously had this dependent on County but that created confounding so I removed that condition.

```{r}
pt <- prop.table(table(dat.sm$OFF_RACER))

dat.sim$OFF_RACER <- factor(sample(1:4, size = N.sim, prob = pt, replace = T),
                            labels = c("WHITE", "BLACK", "LATINO", "OTHER"))

p1 <- my_gg_bar(dat.sm, "OFF_RACER")
p2 <- my_gg_bar(dat.sim, "OFF_RACER")
cowplot::plot_grid(p1, p2, nrow = 1)
```

Check that race is independent of county in the simulated data.
```{r}
race_by_county <- dat.sim %>% group_by(COUNTY) %>%
    summarise(prop_white = mean(OFF_RACER == "WHITE", na.rm = T),
              prop_black = mean(OFF_RACER == "BLACK", na.rm = T),
              prop_latino = mean(OFF_RACER == "LATINO", na.rm = T),
              prop_other = mean(OFF_RACER == "OTHER", na.rm = T)) %>% as.data.frame()
race_by_county %>% arrange(desc(prop_black)) %>% head(10)
race_by_county %>% arrange(desc(prop_white)) %>% head(10)
```


#### YEAR (Categorical, 10 levels)

Indicators for the year from 2010 to 2019
```{r}
pt <- prop.table(table(dat.sm$YEAR))

dat.sim$YEAR <- factor(sample(2010:2019, size = N.sim, prob = pt, replace = T))
p1 <- my_gg_bar(dat.sm, "YEAR")
p2 <- my_gg_bar(dat.sim, "YEAR")
cowplot::plot_grid(p1, p2, nrow = 1)
```


#### Previous Criminal History PRS (Categorical, 4 levels) - 0, 1/2/3, 4/5, REVOC/RFEL

REVOC/RFEL is the most serious
```{r}
pt <- prop.table(table(dat.sm$PRS))

dat.sim$PRS <- factor(sample(1:4, size = N.sim, prob = pt, replace = T), 
                        labels = levels(dat.sm$PRS))
p1 <- my_gg_bar(dat.sm, "PRS")
p2 <- my_gg_bar(dat.sim, "PRS")
cowplot::plot_grid(p1, p2, nrow = 1)
```

#### DOSAGE

Defendant's Age at Sentencing DOSAGE (Numeric, from 11.71 to 113.88)
```{r}
X <- dat.sm$DOSAGE - min(dat.sm$DOSAGE, na.rm = T); N <- nrow(dat.sm)
alpha.hat <- (N - 1) / N *  mean(X, na.rm = T)^2 / var(X, na.rm = T)
beta.hat <- N / (N - 1) * var(X, na.rm = T) / mean(X, na.rm = T)
dat.sim$DOSAGE <- rgamma(N.sim, alpha.hat, scale = beta.hat) + min(dat.sm$DOSAGE, na.rm = T)

p1 <- ggplot(dat.sm, aes(DOSAGE)) +
    geom_density() 
p2 <- ggplot(dat.sim, aes(DOSAGE)) + 
    geom_density()
cowplot::plot_grid(p1, p2, nrow = 1)
```

#### RECMIN

Recommended Minimum Sentence (Binary) - Yes = 1, No = 0
```{r}
pt <- prop.table(table(dat.sm$RECMIN))

dat.sim$RECMIN <- factor(sample(0:1, size = N.sim, prob = pt, replace = T), 
                        labels = levels(dat.sm$RECMIN))
p1 <- my_gg_bar(dat.sm, "RECMIN")
p2 <- my_gg_bar(dat.sim, "RECMIN")
cowplot::plot_grid(p1, p2, nrow = 1)
```

#### Simulating INCAR

Simulating Incarceration decision based on the simulated predictors using the estimated regression coefficients from the CCA logistic regression. Using the regression fit, I assign with a draw from a Bernoulli distribution with the predicted probability for each observation.

```{r}
# creating quadratic terms
dat.sim$OGSQ <- dat.sim$OGS ^ 2
dat.sim$DOSAGEQ <- dat.sim$DOSAGE ^ 2

# Fitted probabilities + some error ~ normal(0, .25^2)
load("full_data_cca_log_reg.rda")
INCAR_p <- predict(fit, dat.sim, type = "response")
rm(fit)

# Outcomes from probs
dat.sim$INCAR <- rbinom(N.sim, 1, INCAR_p)

p1 <- my_gg_bar(dat.sm, "INCAR")
p2 <- my_gg_bar(dat.sim, "INCAR")

cowplot::plot_grid(p1, p2, nrow = 1)
```

```{r}
write_csv(dat.sim, "Data/simulated_data.csv")
```

A quick logistic regression to see how closely these estimates match what I'm taking to be the true estimates from the CCA regression
```{r}
s <- sample(1:nrow(dat.sim), 10000, replace = T)
fit2 <- glm(INCAR ~ ., dat.sim[s, ], family = binomial(link = "logit"))
# summary(fit2)
est.beta <- coef(fit2)

print(paste0("Avg. Bias = ", round(mean(est.beta - beta), 4)))
print(paste0("MSE = ", round(sum((est.beta - beta)^2), 4)))
```

```{r fig.width = 8, fig.height = 4}
coef.names <- names(beta)
bias.df <- data.frame(Covariate = coef.names, 
                      Bias = est.beta - beta)

bias.df$Large_Bias <- ifelse(abs(bias.df$Bias) > 1.5, "Large", "Small/Moderate")
ggplot(bias.df, aes(x = Covariate, y = Bias, color = Large_Bias)) +
    geom_point() +
    geom_text(aes(label = ifelse(Large_Bias == "Large", Covariate, "")),
              nudge_y = 1) +
    geom_hline(yintercept = c(0, 1.5, -1.5)) +
    coord_cartesian(xlim = c(-1, 93)) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
          legend.position = "bottom")
```




