---
title: "Criminology Simulation Results"
author: "Ben Stockton"
number-sections: true 
df-print: paged
pdf-engine: pdflatex
toc: true
format: 
  html: 
    code-fold: false
  pdf:
    geometry: 
      - top=30mm
      - left=30mm
    shift-heading-level-by: -1
  docx:
    shift-heading-level-by: -1
editor: visual
fig-width: 7
fig-height: 5
fig-dpi: 300
keep-tex: true
<<<<<<< Updated upstream
=======
keep-md: true
>>>>>>> Stashed changes
execute: 
  cache: true
bibliography: ../../Literature/Criminology.bib
---

```{r}
#| include: false
knitr::opts_chunk$set(echo = FALSE, results = 'hold')
library(dplyr, warn.conflicts = FALSE)

# Suppress summarise info
options(dplyr.summarise.inform = FALSE)
library(magrittr)
library(ggplot2)
library(latex2exp)
library(RColorBrewer)
library(forcats)
source("../../R/helpers.R")
source("../../R/results_figures.R")
source("../../R/results_tables.R")
```

## Overview

I ran the simulations under a variety of settings, each for 1000 iterations.

| Sample Size ($n$) | Number of Imputations ($M$) |
|-------------------|-----------------------------|
| 500               | 3                           |
| 500               | 5                           |
| 500               | 8                           |
| 1000              | 3                           |
| 1000              | 5                           |
| 1000              | 8                           |
| 2500              | 3                           |
| 2500              | 5                           |
| 2500              | 8                           |

From the estimated rate of missing information, I determined that there should be between five and eight imputations to achieve a relative efficiency of \>0.99. For a relative efficiency of \~0.95, three imputations is sufficient.

In previous simulations with $n = 1000$ and $M = 3$ , I was able to find that on average, depending on my parameter settings for the missingness model, I was able to create data sets that over- and under-estimated the race effect on average for each data set. In other words, for a given data set, if you take a sample from the population with race effect $\gamma = \exp(\beta)$ and estimate the race effect with the complete data to get $\hat{\gamma} = \exp(\hat{\beta})$, then estimates made with complete case analysis after imposing missing values for that same data set will be over or under the complete data estimate $\hat{\gamma}$ on average, depending on what parameters are used in the missingness model.

In this document, I will demonstrate that for appropriate numbers of imputations to achieve 95% relative efficiency, 99% relative efficiency, and 99+% relative efficiency, we see the same behavior in the sampling distributions and medians of the differences $\hat{\gamma}^{miss} - \hat{\gamma}$.

## Results

### n = 500

```{r}
#| label: "n = 500 Results Tables"
#| echo: false
Q <- 1000
N <- c(500, 1000, 2500)
m <- c(3, 5, 8)

tbs <- present_tables(Q = Q, N = N[1], m = m)

plts <- result_plots(tbs$sim_data, Q = Q, N = N[1], horizontal = T)

print(plts$DIFF)
tbs$tables$DIFF |> 
    filter(`Miss Type` == "MAR") |> 
    kableExtra::kbl(format = "latex", digits = 3, booktabs = TRUE)
tbs$tables$PROP_MISS |> 
    filter(`Miss Type` == "MAR") |> 
    kableExtra::kbl(format = "latex", digits = 3, booktabs = TRUE)
print(plts$BIAS)
tbs$tables$BIAS |> 
    filter(`Miss Type` == "MAR") |> 
    kableExtra::kbl(format = "latex", digits = 3, booktabs = TRUE)
```

### n = 1000

```{r}
#| label: "n = 1000 Results Tables"
#| echo: false

tbs <- present_tables(Q = Q, N = N[2], m = m)

plts <- result_plots(tbs$sim_data, Q = Q, N = N[2], horizontal = T)

print(plts$DIFF)
tbs$tables$DIFF |> kableExtra::kbl(format = "latex", digits = 3, booktabs = TRUE)
tbs$tables$PROP_MISS |> kableExtra::kbl(format = "latex", digits = 3, booktabs = TRUE)
print(plts$BIAS)
tbs$tables$BIAS |> kableExtra::kbl(format = "latex", digits = 3, booktabs = TRUE)
```

### n = 2500

```{r}
#| label: "n = 2500 Results Tables"
#| echo: false

tbs <- present_tables(Q = Q, N = N[3], m = m)

plts <- result_plots(tbs$sim_data, Q = Q, N = N[3], horizontal = T)

print(plts$DIFF)
tbs$tables$DIFF |> kableExtra::kbl(format = "latex", digits = 3, booktabs = TRUE)
tbs$tables$PROP_MISS |> kableExtra::kbl(format = "latex", digits = 3, booktabs = TRUE)
print(plts$BIAS)
tbs$tables$BIAS |> kableExtra::kbl(format = "latex", digits = 3, booktabs = TRUE)
```

## MAR Overall Results Table

```{r}
sim_settings <- readr::read_csv("../sim_settings.csv")
sim_settings[sim_settings$n < 5000, "simQ"] <- 1000

sim_settings <- sim_settings[-c(22, 23, 24, 35, 36),]

sim_res_list <- list()

for (i in 1:nrow(sim_settings)) {
    Q <- sim_settings[i, "simQ"]
    n <- sim_settings[i, "n"]
    m <- sim_settings[i, "M"]
    fname <- paste0("../simulation_results_Q", Q, "_n_", n, "_m_", m, "_p_miss_0.01.csv")
    tmp <- read.csv(fname, header = T)
    tmp$IMPUTATION <- paste0(m, " imps")
    tmp$N <- as.numeric(n)
    sim_res_list[[i]] <- tmp
    
}

sim_res_list %>% purrr::reduce(bind_rows) -> sim_res

sim_res[which(sim_res$DIRECTION == "COMP"), "ANALYSIS"] <- "COMP"
sim_res[which(sim_res$DIRECTION == "COMP"), "DIRECTION"] <- "No Missing Data"
sim_res$DIRECTION %<>% fct_relevel(c("OVER", "UNDR", "No Missing Data"))
sim_res$ANALYSIS %<>% fct_relevel(c("CCA", "COMP", "MI"))
sim_res$IMPUTATION %<>% fct_relevel(paste0(c(3, 5, 8, 25, 50, 100), " imps"))
```

```{r}
options(dplyr.summarise.inform = FALSE)
sim_res |> 
  filter(IMPUTATION == "3 imps", 
         MISS_TYPE == "MAR") |>
  group_by(N, DIRECTION, ANALYSIS) |>
  summarize(
    OR_DIFF = mean(OR_DIFF),
    OR_BIAS = mean(OR_BIAS),
    PROP_MISS = mean(PROP_MISS)
  ) |> 
  kableExtra::kbl(format = "latex", 
                  digits = 3, 
                  booktabs = TRUE,
                  caption = "MAR simulation results with 3 imputations for MI.")

sim_res |> 
  filter(IMPUTATION == "5 imps", 
         MISS_TYPE == "MAR") |>
  group_by(N, DIRECTION, ANALYSIS) |>
  summarize(
    OR_DIFF = mean(OR_DIFF),
    OR_BIAS = mean(OR_BIAS),
    PROP_MISS = mean(PROP_MISS)
  ) |> 
  kableExtra::kbl(format = "latex", 
                  digits = 3, 
                  booktabs = TRUE,
                  caption = "MAR simulation results with 5 imputations for MI.")

sim_res |> 
  filter(IMPUTATION == "8 imps", 
         MISS_TYPE == "MAR") |>
  group_by(N, DIRECTION, ANALYSIS) |>
  summarize(
    OR_DIFF = mean(OR_DIFF),
    OR_BIAS = mean(OR_BIAS),
    PROP_MISS = mean(PROP_MISS)
  ) |> 
  kableExtra::kbl(format = "latex", 
                  digits = 3, 
                  booktabs = TRUE,
                  caption = "MAR simulation results with 8 imputations for MI.")

```

```{r}
#| label: fig-mar-diff-pmiss-001
#| fig-cap: "Distributions of the difference between the complete data estimate and the incomplete data estimates under complete case analysis (CCA), or multiple imputation (MI) with 3, 5, 8, 25, 50, and 100 imputations (imps) from simulations at sample sizes from N = 500 to N = 50000."
sim_res |> 
  filter(MISS_TYPE == "MAR",
         ANALYSIS != "COMP") |>
  mutate(
    ANALYSIS = case_when(
      ANALYSIS == "MI" ~ paste0("MI ", IMPUTATION),
      ANALYSIS == "CCA" ~ "CCA"
    )
  ) -> sim_res_sum
sim_res_sum$ANALYSIS %<>% fct_relevel(c("CCA", paste0("MI ", c(3, 5, 8, 25, 50, 100), " imps")))
sim_res_sum |>
  ggplot(aes(DIRECTION, OR_DIFF, color = as.factor(N))) +
        geom_hline(yintercept = 0, color = "gray") +
        geom_boxplot(fill = "white", alpha = .75, outlier.shape = 1) +
        facet_wrap(ANALYSIS ~ .) +
        theme_bw() +
        ggthemes::scale_fill_colorblind() +
        ggthemes::scale_color_colorblind(name = 'Sample Size') +
        theme(legend.position = "right") +
        labs(title = "Difference in Race Effect Estimates \n from CCA/MI to the Complete Data Estimate",
             x = "Intended Direction of Bias due to Missing Data",
             y = "Difference in the OR")
```

```{r}
#| label: fig-mar-bias-pmiss-001
#| fig-cap: "Distributions of the statistical bias between the true race effect parameter and the estimates from the complete data (COMP), complete case analysis (CCA) estimate, and multiple imputation (MI) with 3, 5, 8, 25, 50, and 100 imputations (imps) from simulations at sample sizes from N = 500 to N = 50000."
sim_res |> 
  filter(MISS_TYPE == "MAR") |>
  mutate(
    ANALYSIS = case_when(
      ANALYSIS == "MI" ~ paste0("MI ", IMPUTATION),
      ANALYSIS == "CCA" ~ "CCA",
      ANALYSIS == "COMP" ~ "COMP"
    )
  ) -> sim_res_sum
sim_res_sum$ANALYSIS %<>% fct_relevel(c("COMP", "CCA", paste0("MI ", c(3, 5, 8, 25, 50, 100), " imps")))
sim_res_sum |>
  ggplot(aes(DIRECTION, OR_BIAS, color = as.factor(N))) +
        geom_hline(yintercept = 0, color = "gray") +
        geom_boxplot(fill = "white", alpha = .75, outlier.shape = 1) +
        facet_wrap(. ~ ANALYSIS, scales = "free_x") +
        theme_bw() +
        ggthemes::scale_fill_colorblind() +
        ggthemes::scale_color_colorblind(name = 'Sample Size') +
        theme(legend.position = "right") +
        labs(title = "Statistical Bias of the Race Effect Estimates", 
             x = "Intended Direction of Bias due to Missing Data",
             y = "Statistical Bias")
```

