---
title: "Criminology Simulation Results"
author: "Ben Stockton"
number-sections: true 
df-print: paged
pdf-engine: pdflatex
toc: true
format: 
  html: 
    code-fold: false
  pdf:
    geometry: 
      - top=30mm
      - left=30mm
    shift-heading-level-by: -1
  docx:
    shift-heading-level-by: -1
editor: visual
execute: 
  cache: true
bibliography: ../../Literature/Criminology.bib
---

```{r}
#| include: false
knitr::opts_chunk$set(echo = TRUE, results = 'hold')
library(dplyr, warn.conflicts = FALSE)

# Suppress summarise info
options(dplyr.summarise.inform = FALSE)
library(magrittr)
library(ggplot2)
library(latex2exp)
library(RColorBrewer)
library(forcats)
source("../../R/helpers.R")
source("../../R/results_figures.R")
source("../../R/results_tables.R")
```

## Overview

I ran the simulations under a variety of settings, each for 225 iterations.

| Sample Size ($n$) | Number of Imputations ($M$) |
|-------------------|-----------------------------|
| 500               | 3                           |
| 500               | 5                           |
| 500               | 8                           |
| 1000              | 3                           |
| 1000              | 5                           |
| 1000              | 8                           |
| 2500              | 3                           |
| 2500              | 5                           |
| 2500              | 8                           |

From the estimated rate of missing information, I determined that there should be between five and eight imputations to achieve a relative efficiency of \>0.99. For a relative efficiency of \~0.95, three imputations is sufficient.

In previous simulations with $n = 1000$ and $M = 3$ , I was able to find that on average, depending on my parameter settings for the missingness model, I was able to create data sets that over- and under-estimated the race effect on average for each data set. In other words, for a given data set, if you take a sample from the population with race effect $\gamma = \exp(\beta)$ and estimate the race effect with the complete data to get $\hat{\gamma} = \exp(\hat{\beta})$, then estimates made with complete case analysis after imposing missing values for that same data set will be over or under the complete data estimate $\hat{\gamma}$ on average, depending on what parameters are used in the missingness model.

In this document, I will demonstrate that for appropriate numbers of imputations to achieve 95% relative efficiency, 99% relative efficiency, and 99+% relative efficiency, we see the same behavior in the sampling distributions and medians of the differences $\hat{\gamma}^{miss} - \hat{\gamma}$.

## Results

### n = 500

```{r}
#| label: "n = 500 Results Tables"
#| echo: false
Q <- 1000
N <- c(500, 1000, 2500)
m <- c(3, 5, 8)

tbs <- present_tables(Q = Q, N = N[1], m = m)

plts <- result_plots(tbs$sim_data, Q = 225, N = N[1], horizontal = T)

print(plts$DIFF)
print(tbs$tables$DIFF)
print(plts$BIAS)
```

### n = 1000

```{r}
#| label: "n = 1000 Results Tables"
#| echo: false

tbs <- present_tables(Q = Q, N = N[2], m = m)

plts <- result_plots(tbs$sim_data, Q = 225, N = N[2], horizontal = T)

print(plts$DIFF)
print(tbs$tables$DIFF)
print(plts$BIAS)
```

### n = 2500

```{r}
#| label: "n = 2500 Results Tables"
#| echo: false

tbs <- present_tables(Q = Q, N = N[3], m = m)

plts <- result_plots(tbs$sim_data, Q = 225, N = N[3], horizontal = T)

print(plts$DIFF)
print(tbs$tables$DIFF)
print(plts$BIAS)
```
